Carbonate chemistry at Ny-Ålesund: AWIPEV-CO~2~ Project

### 2020-09-03 awipev-CO2.Rmd
consideration on stratification in the air-sea CO2 flux section

```{r seafet profile, echo=FALSE, message= FALSE, warning=FALSE, out.width="80%",fig.show='hold', fig.align="center"}
# pH by depth
# d_all<- d_all%>%
#   dplyr::mutate(year = year(datetime))

#Per month
range <- d_all %>%
  dplyr::mutate(pH_range_more_6m = ifelse(pressure_insitu_ctd > 6, pHint_tot_sf, NA)) %>%
  group_by(montht) %>%
  summarize(pH_range = max(pHint_tot_sf, na.rm = TRUE) - min(pHint_tot_sf, na.rm = TRUE),
            pH_range_more_6m = max(pH_range_more_6m, na.rm = TRUE) - min(pH_range_more_6m, na.rm = TRUE),
            pH_range_more_9m = max(pHint_tot_sf_9m, na.rm = TRUE) - min(pHint_tot_sf_9m, na.rm = TRUE)) %>%
  ungroup()
print(kable(range, digits = 2))

tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf, depth) %>%
  group_by(depth, montht) %>%
  ungroup()
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth), y = pHint_tot_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pHT", x = "Depth (m)", y= "pHT") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

```
**Conclusion**:
The range is significantly larger during some months when the data deeper than 9 m are considered. Also, pH is sometimes higher at the surface, at other tilmes it is higher on the bottom (June). 
  + To compare with pCO2/pCO2 measured in the Ferryboxt seems therefore better to stick with pH_9m, that is at depths deeper than 8 m.
  + For air-sea CO2 fluxes

@@

To calculate the air-sea CO2 flux, we use:
+ seawater pCO2
+ atmospheric CO2 measured at the Zeppelin station and downloaded 2020-08-19 from https://gaw.kishou.go.jp/search/file/0054-6001-1001-01-01-9999 
+ wind spreed data collected by the AWI at 10 m height (sent to JPG by P. Fischer, Aug. 2020)

@@

>>>>

```{r seafet profile, echo=FALSE, message= FALSE, warning=FALSE, out.width="80%",fig.show='hold', fig.align="center"}
# range <- d_all %>%
#   dplyr::mutate(pH_range_more_6m = ifelse(pressure_insitu_ctd > 6, pHint_tot_sf, NA)) %>%
#   group_by(montht) %>%
#   summarize(pH_range = max(pHint_tot_sf, na.rm = TRUE) - min(pHint_tot_sf, na.rm = TRUE),
#             pH_range_more_6m = max(pH_range_more_6m, na.rm = TRUE) - min(pH_range_more_6m, na.rm = TRUE),
#             pH_range_more_9m = max(pHint_tot_sf_9m, na.rm = TRUE) - min(pHint_tot_sf_9m, na.rm = TRUE)) %>%
#   ungroup()
# print(kable(range, digits = 2))

# # datetime when pH values are in but no pressure data. File sent to send to Philipp
# no_pressure <- d_all %>% 
#   dplyr::select(datetime, pressure_insitu_ctd, pHint_tot_sf) %>% 
#   dplyr::filter(is.na(pressure_insitu_ctd) & !is.na(pHint_tot_sf))
# readr::write_csv(x = no_pressure, path=paste0(path,"/fb_awipev-co2_server/ny-alesund/data//no_pressure.csv"))

# we look at all depths
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf, depth) %>%
  group_by(depth, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth), y = pHint_tot_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pHT", x = "Depth (m)", y= "pHT") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

# we look at 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf, depth) %>%
  dplyr::mutate(depth = as.factor(ifelse(depth == 1 | depth == 3, 2, depth))) %>% 
  dplyr::filter(depth == 2 | depth == 9) %>% 
  group_by(depth, montht)
summary(aov(tmp$pHint_tot_sf ~ tmp$depth * tmp$montht))

p <- tmp %>% 
  ggplot(aes(x=as.factor(depth), y = pHint_tot_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pHT", x = "Depth (m)", y= "pHT") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
```
**Conclusion**:
The range is significantly larger during some months when the data deeper than 8 m are considered. Also, pH is sometimes higher at the surface, at other times it is higher on the bottom (June). No data are available at 1 m (actually 0 to 2 m). So, to compare surface and bottom, I combine 1 and 3 m (0 to 4 m) and 9 m (below 8 m).

Depth and month of the year have a statistically significant effet on pH. The interaxction is also significant, indicating that the effect of depth differs depending on the month.

 + To compare with pCO2/pCO2 measured in the Ferrybox seems therefore better to stick with pH_9m, that is at depths deeper than 8 m.
  + pCO2 would have the same behaviour as pH. For air-sea CO2 fluxes, one should use data from 0 to 4 m.

@@

To calculate the air-sea CO2 flux, we use:

+ seawater pCO2 measured in the FerryBox on water pumped from 11 m depth.
+ atmospheric CO2 measured at the Zeppelin station and downloaded on 2020-08-19 from https://gaw.kishou.go.jp/search/file/0054-6001-1001-01-01-9999 
+ wind spreed data collected by the AWI at 10 m height (sent to JPG by P. Fischer, Aug. 2020)

@@

seaFET pH is significantly different at the surface and depth. It should therefore be the same for pCO2.

Let's look first at the stratification (salinity, temperature and density).

```{r stratification, echo=FALSE, message= FALSE, warning=FALSE}
# we look at salinity all depths
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, depth) %>%
  group_by(depth, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth), y = sal_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Salinity vs depth", x = "Depth (m)", y= "Salinity") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

# we look at salinity 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, temp_insitu_ctd, depth) %>%
  dplyr::mutate(depth = as.factor(ifelse(depth == 1 | depth == 3, 2, depth))) %>% 
  dplyr::filter(depth == 2 | depth == 9) %>% 
  group_by(depth, montht)
summary(aov(tmp$temp_insitu_ctd ~ tmp$depth * tmp$montht))

p <- tmp %>% 
  ggplot(aes(x=as.factor(depth), y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Salinity vs depth", x = "Depth (m)", y= "Salinity") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

# we look at temperature all depths
tmp <- d_all %>% 
  dplyr::select(montht, temp_insitu_ctd, depth) %>%
  group_by(depth, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth), y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Temperature vs depth", x = "Depth (m)", y= "Salinity") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

# we look at temperature 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, temp_insitu_ctd, depth) %>%
  dplyr::mutate(depth = as.factor(ifelse(depth == 1 | depth == 3, 2, depth))) %>% 
  dplyr::filter(depth == 2 | depth == 9) %>% 
  group_by(depth, montht)
summary(aov(tmp$temp_insitu_ctd ~ tmp$depth * tmp$montht))

p <- tmp %>% 
  ggplot(aes(x=as.factor(depth), y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Temperature vs depth", x = "Depth (m)", y= "Salinity") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

# we look at density 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, temp_insitu_ctd, depth) %>%
  dplyr::mutate(density = seacarb::rho(S = sal_insitu_ctd, T = temp_insitu_ctd, P = 0),
                depth = as.factor(ifelse(depth == 1 | depth == 3, 2, depth))) %>% 
  dplyr::filter(depth == 2 | depth == 9) %>% 
  group_by(depth, montht)
summary(aov(tmp$density ~ tmp$depth * tmp$montht))

p <- tmp %>% 
  ggplot(aes(x=as.factor(depth), y = density)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Density vs depth", x = "Depth (m)", y= "Salinity") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
```
